Rule Synthesis for Rule-based Intrusion Detectors

### 2020-05-19 paper/paper.tex
pass
pass on Step 2 (RUles Creation)

\Mar{--------------------------------------------- I am here}

@@

To address the overfiting issue, \tname{} leverages public data sets
of benign traffic~\cite{tcpreplay} and additional examples of the
attack to guide the search for more general rules. To that end,
\tname{} \emph{minimizes} the rules produced in the previous step,
preserving the invariant of \emph{only} capturing the positive
malicious traffic. We say that a rule is minimal if discarding any
option results in the capture of negative (\ie{}, benign) traffic.
The procedure we propose in this step builds on the observation that
all options from the rule produced in the previous step are satisifed
by the traffic produced by the attack. Consequently, any subset of
these options also captures the attack. The rule creation procedure
iteratively discards options from visited rules---derived from the
seed rule---using the negative traffic to guide the search for minimal
rules.

@@

Algorithm~\ref{algo1} shows the pseudocode for deriving rules from the
seed rule produced at the Reverse Engineering step. The algorithm
takes as input an over-constrained rule $sr$ and pcap files~\cite{pcap}
to reconstruct the benign ($\mathit{bTr}$) and malicious traffic
($\mathit{mTr}$). The algorithm produces on output a set of
alternative rules to $sr$ that captures every instance of the attack
(as per $\mathit{mTr}$), but does not capture any message from the
benign traffic (as per $\mathit{bTr}$).

Line~\ref{line-init} initializes the current population of individuals
(\ie{}, rules) $\mathcal{R}$ with the seed rule $sr$. We encode a rule
as a bitvector where each index in the vector represents one option;
the value assigned at a given position indicates whether or not the
option is present in the rule. One iteration of the outer loop
(lines~\ref{start-do-while}-\ref{end-do-while}) processes the current
population of individuals $\mathcal R$ and creates the next generation
$\mathcal R'$. Line~\ref{init-rprime} initializes the population of
individuals for the next generation. The inner loop
(lines~\ref{start-foreach}--\ref{end-foreach}) iterates through each
candidate rule in $\mathcal R$. The algorithm mutates each of these
rules for $N$ times, discarding one option of the rule each
time. Line~\ref{copy} assigns a copy of the bitvector, encoding
$\mathit{ro}$, to variable $r$. Function $\mathit{len}$ returns the
length of a vector and function $\mathit{rand}$ randomly chooses one
integer value in the range defined by its parameters. So, the
expression $\mathit{rand(0, len(r))}$, appearing at
line~\ref{choose-index}, returns an index of the option in $r$ to
delete at line~\ref{delete-option}. The call to $\mathit{isNew}$
checks if $r$ has not been previously generated in another iteration
whereas the call to $\mathit{isPlausible}$ checks if the rule only
captures positive traffic, \ie{}, it checks that no message present in
$\mathit{bTr}$ is captured by $r$. If $r$ satisfies both criteria, it
is added to $\mathcal R'$, which denotes the next generation of
individuals in the search. After processing every candidate rule in
$\mathcal R$, the next generation of individuals is defined
(line~\ref{define-next-gen}) and the set of discovered rules is
updated (at line~\ref{output-accum}).

@@

Note that the algorithm produces no rules in the last iteration. Rules
with a single option are produced in the second-to-last iteration. In
other iterations, rules with more than one option are
produced. Conceptually, the search induces a search tree with the seed
rule $sr$ as root, single-option rules as child nodes, and rules with
more options as internal nodes. The edges of the tree denote the
ancestor relation between a pair of rules. It is also worth noting
that the algorithm reports every rule in the tree instead of only
those with one option. The reason is that the traffic used to check
plausibility is limited by definition---reporting only minimal rules
with respect to that traffic could discard the correct rule.

In summary, the algorithm reports many rules that are plausible by
using the benign traffic to reduce the over-constrained rule produced
at the \reveng\ step. In the Raking step (Section~\ref{sec:ranking}),
\tname{} ranks the rules this algorithm reports by their similarity to
trusted rules. Intuitively, rules dissimilar to those trusted rules
rank low.

@@

\textbf{Implementation details.} We implemented function
\emph{isPlausible} by 1)~spawning Suricata to monitor a specified
rule, 2)~replaying the negative benign traffic (see
Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
check if any benign message was captured by the rule. The function
\emph{filter} is implemented similarly. It returns a projection of the
set $\mathcal O$ where every rule captures the attacks characterized
by $\mathit{mTr}$. It is worth noting that Suricata can process
\emph{pcap} files~\cite{pcap} very efficiently compared to analyzing
network messages directly. Instead of sending traffic to a machine
with a deployed server (\eg{}, web server) that \suri\ would monitor,
our setup calls \suri\ directly through an API. \suri\ is invoked
multiple times for creating rules.

@@

@@

(RCE)~\cite{remote-code-execution}, and Local File Inclusion
(LFI)~\cite{local-file-inclusion}.

>>>>

@@

To address the overfiting issue, \tname{} leverages public data sets
of benign traffic~\cite{tcpreplay} and additional examples of the
attack to guide the search for more general rules. To that end,
\tname{} \emph{minimizes} the rules produced in the previous step,
preserving the invariant of \emph{only} capturing the positive
malicious traffic. We say that a rule is minimal if discarding any
option results in the capture of negative (\ie{}, benign) traffic.
The procedure we propose in this step builds on the observation that
all options from the rule produced in the previous step are satisifed
by the traffic produced by the attack. Consequently, any subset of
these options also captures the attack.

@@

The rule creation procedure \tname{} uses iteratively discards options
from rules, derived from the seed rule, using the negative traffic to
guide the search for minimal rules.  Algorithm~\ref{algo1} shows the
pseudocode for deriving these rules. The algorithm takes as input an
over-constrained rule $sr$, denoting the seed rule produced at the
Reverse Engineering step, and pcap files~\cite{pcap} to reconstruct
the benign traffic ($\mathit{bTr}$) and the malicious traffic
($\mathit{mTr}$). The algorithm produces on output a set of
alternative rules to $sr$ that captures every instance of the attack
(as per $\mathit{mTr}$), but does not capture any message from the
benign traffic (as per $\mathit{bTr}$).

Line~\ref{line-init} initializes the current population of individuals
(\ie{}, rules) $\mathcal{R}$ with the seed rule $sr$. We encode a rule
as a bitvector where each index in the vector represents one option;
the value assigned at a given position indicates whether or not the
option is present in the rule. One iteration of the outer loop
(lines~\ref{start-do-while}-\ref{end-do-while}) processes the current
population of individuals $\mathcal R$ and creates the next generation
$\mathcal R'$. Line~\ref{init-rprime} initializes the population of
individuals for the next generation. The inner loop
(lines~\ref{start-foreach}--\ref{end-foreach}) iterates through each
candidate rule in $\mathcal R$. The algorithm mutates each of these
rules for $N$ times, discarding one rule option each
time. Line~\ref{copy} assigns a copy of the bitvector encoding
$\mathit{ro}$ to variable $r$. Function $\mathit{len}$ returns the
length of a vector and function $\mathit{rand(lo, hi)}$ randomly
chooses one integer value in the range $[lo, hi]$ defined by its
parameters. So, the expression $\mathit{rand(0, len(r))}$, appearing
at line~\ref{choose-index}, returns an index of the option in $r$ and
the statement at line~\ref{delete-option} deletes that option. The
call to $\mathit{isNew}$ checks if $r$ has not been previously
generated in another iteration whereas the call to
$\mathit{isPlausible}$ checks if the rule only captures positive
traffic, \ie{}, it checks that no message present in $\mathit{bTr}$ is
captured by $r$. If $r$ satisfies both criteria, it is added to
$\mathcal R'$, which denotes the next generation of individuals in the
search. After processing every candidate rule in $\mathcal R$, the
next generation of individuals is defined (line~\ref{define-next-gen})
and the set of discovered rules is updated (at
line~\ref{output-accum}).

@@

Note that the algorithm produces no rules in the last iteration. Rules
with a single option are produced in the second-to-last iteration. In
other iterations, rules with more than one option are produced. The
search induces a search tree with the seed rule $sr$ as root,
single-option rules as child nodes, and rules with more options as
internal nodes. The edges of the tree denote the ancestor relation
between a pair of rules. It is also worth noting that the algorithm
reports every rule in the tree instead of only those with one
option. The reason is that the traffic used to check plausibility is
limited by definition---reporting only minimal rules with respect to
that traffic could discard the correct rule.

In summary, the algorithm reports many rules that are plausible by
using the benign traffic to reduce the over-constrained rule produced
at the \reveng\ step. In the Raking step (Section~\ref{sec:ranking}),
\tname{} ranks the rules.

%Intuitively, rules dissimilar to (a very
%large set of) trusted rules rank low.

% this algorithm reports by their similarity to trusted rules

@@

\textbf{Implementation details.} We implemented function
\emph{isPlausible} by 1)~spawning Suricata to monitor a specified
rule, 2)~replaying the benign traffic (see
Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
check if any benign message was captured by the rule. The function
\emph{filter} is implemented similarly. It returns a projection of the
set $\mathcal O$ where every rule captures the attacks characterized
by $\mathit{mTr}$. It is worth noting that Suricata can process
\emph{pcap} files~\cite{pcap} very efficiently compared to analyzing
network messages directly. Instead of sending traffic to a machine
with a deployed server (\eg{}, web server) that \suri\ would monitor,
our setup calls \suri\ directly through an API. \suri\ is invoked
multiple times for creating rules.

@@

\Mar{-------------------------- I am here}

@@

(RCE)~\cite{remote-code-execution}, Local File Inclusion
(LFI)~\cite{local-file-inclusion} and ICMP Traffic Alerts \cite{icmp-types}.
